FoRTE baseline

### 2019-06-06 analysis/paper/paper.Rmd
Update references
Render ED2 list in intro correctly

Instead, more reliable predictions can be obtained by using dynamic vegetation models that explicitly represent process involved in forest growth and mortality.
Vegetation models fall broadly along the following spectrum of complexity.
On one side are relatively simple "Big leaf" models---such as PNET, SiBCASA, and Biome-BGC <!-- TODO refs; more examples? -->---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the (weighted) average of all the vegetation at that site. <!-- TODO refs -->
These models were designed originally to simulate the land surface boundary condition for atmospheric general circulation models, though similar models continue to be applied successfully to simulate biogeochemical processes in many ecosystem science contexts. <!-- TODO refs -->
On the other end of the spectrum are "individual-based" models (a.k.a. "gap models"), which explicitly simulate multiple indivduals competing for resources at a single site [@shugart_2015_computer]. <!-- TODO other refs? -->
Examples of such models are LANDIS <!-- TODO ref !--> and UVAFME. <!-- TODO ref --> <!-- TODO Others? -->
Because these models can explicitly represent inter-specific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive].
Moreover, these models' predictions of species composition and ecosystem structure are relevant outside of carbon cycle science, particularly for forest management and wildlife conservation. <!-- TODO refs -->
However, these benefits of individual-based models come at the cost of much higher data and computational requirements, which effectively restrict their application to individual, well-studied sites. <!-- TODO refs. Also, is this actually true? -->

A key challenge of vegetation modeling research over the last several decades has been the development of approaches that capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual [@purves_2008_predictive].
One such approach is the "ecosystem demography" approach, which uses a system of partial differential equations to simulate the emergent mean behavior of "cohorts" of individuals of similar composition, size, and age [@moorcroft_2001_method]. <!-- TODO Run this description by Mike? -->
The original "Ecosystem Demography" (ED) model [@moorcroft_2001_method] was enhanced by @medvigy_2009_mechanistic (as ED2) with fully-coupled leaf physiology, canopy radiative transfer, and micrometeorology, and each of these components have been further refined over the last decade [culminating in ED2.2; @longo_2019_ed1].
Various versions of the ED model have been validated in boreal [@medvigy_2011_predicting], temperate [@medvigy_2009_mechanistic; @dietze_2014_quantitative; @raczka_2018_what], and tropical biomes [@moorcroft_2001_method; @longo_2019_ed2], and have been applied in:
- paleoecological studies [@rollinson_2017_emergent];
- free-air CO~2~ enrichment studies [@dekauwe_2013_forest; @miller_2015_alteration];
- estimation of potential carbon stocks [@hurtt_2004_beyond_potential_vegetation]
- vegetation-climate feedbacks {Swann, etc} <!-- TODO -->

>>>>

Instead, more reliable predictions can be obtained by using dynamic vegetation models that explicitly represent process involved in forest growth and mortality.
Vegetation models fall broadly along the following spectrum of complexity.
On one side are relatively simple "Big leaf" models---such as PNET, SiBCASA, and Biome-BGC <!-- TODO refs; more examples? -->---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the (weighted) average of all the vegetation at that site. <!-- TODO refs -->
These models were designed originally to simulate the land surface boundary condition for atmospheric general circulation models, though similar models continue to be applied successfully to simulate biogeochemical processes in many ecosystem science contexts. <!-- TODO refs -->
On the other end of the spectrum are "individual-based" models (a.k.a. "gap models"), which explicitly simulate multiple indivduals competing for resources at a single site [@shugart_2015_computer]. <!-- TODO other refs? -->
Examples of such models are LANDIS <!-- TODO ref !--> and UVAFME. <!-- TODO ref --> <!-- TODO Other examples? -->
Because these models can explicitly represent inter-specific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive].
Moreover, these models' predictions of species composition and ecosystem structure are relevant outside of carbon cycle science, particularly for forest management and wildlife conservation. <!-- TODO refs -->
However, these benefits of individual-based models come at the cost of much higher data and computational requirements, which effectively restrict their application to individual, well-studied sites. <!-- TODO refs. Also, is this actually true? -->

A key challenge of vegetation modeling research over the last several decades has been the development of approaches that capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual [@purves_2008_predictive].
One such approach is the "ecosystem demography" approach, which uses a system of partial differential equations to simulate the emergent mean behavior of "cohorts" of individuals of similar composition, size, and age [@moorcroft_2001_method]. <!-- TODO Run this description by Mike? -->
The original "Ecosystem Demography" (ED) model [@moorcroft_2001_method] was enhanced by @medvigy_2009_mechanistic (as ED2) with fully-coupled leaf physiology, canopy radiative transfer, and micrometeorology, and each of these components have been further refined over the last decade [culminating in ED2.2; @longo_2019_ed1].
Various versions of the ED model have been validated in boreal [@medvigy_2011_predicting], temperate [@medvigy_2009_mechanistic; @dietze_2014_quantitative; @raczka_2018_what], and tropical biomes [@moorcroft_2001_method; @longo_2019_ed2], and have been applied in:

- paleoecological studies [@rollinson_2017_emergent];
- free-air CO~2~ enrichment studies [@dekauwe_2013_forest; @miller_2015_alteration];
- estimation of potential carbon stocks [@hurtt_2004_beyond_potential_vegetation]
- vegetation-climate feedbacks {Swann, etc} <!-- TODO -->
